A Spatial AR System for Wide-area Axis-aligned Metric Augmentation of Planar Scenes

### 2021-10-09 paper/elsarticle/elsarticle-template.tex
paper

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to such instructions in paper form or on a statically situated screen. \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for any to hold a handheld device such as a tablet or for each to wear some form of head-mounted display. In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer. For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that is cumbersome and time consuming. Another drawback of conventional spatial AR relying only on a projector is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.

We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror to direct the projection across varying target locations and a camera facing the scene plane in support of calibrating the system. Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways. First, (i) placement of the desired augmentations is carried out in accordance with the axes of an image of the scene acquired by the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene. Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations. %Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
\end{abstract}

@@

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
\begin{equation}
\mathtt{K} = \begin{bmatrix}
f & 0 & x_0 \\
0 & f & y_0 \\
0 & 0 & 1
\end{bmatrix},
\end{equation}
where $f$ and $(p_x, p_y)^\top$ are both expressed in units of pixels. The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the appropriate lens distortion model coefficients.

@@

\begin{figure}
    \centering
    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
    \qquad
    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
    \caption{Recovering 2D positions in support of projector calibration. (a) 2D positions of the 2D-3D correspondences to be used for calibrating the projector are obtained by detecting---in the image plane of the projector---the circle centers in the asymmetrical circle pattern image. It is this pattern that we project to each of the target locations in the scene plane. (b) For each such target location, an image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected, in the image plane of the camera. A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected. Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
    \label{fig:2d}
\end{figure}

\paragraph{Camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.

\paragraph{Projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences, yet for projector calibration we obtain them by \textit{projecting} a calibration pattern that we detect using the calibrated downwards-facing camera. The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector. We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each. Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}). Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

\begin{figure}
    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane (red). Note that as in the figures of this section that follow, the rendering in the figure corresponds to the projection calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
    \label{fig:3d}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3.png}}}
    \caption{The virtual camera is obtained by (i) rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane. The virtual projector is obtained by (ii) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (iii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iv) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out---likewise intuitively---with respect to the ground plane.

>>>>

\begin{abstract}
Augmented reality (AR) promises to enable use cases in industrial settings that include the embedding of assembly instructions directly into the scene, potentially reducing or altogether obviating the need for workers to refer to such instructions in paper form or on a statically situated screen. \textit{Spatial} AR, in turn, is a form of AR whereby the augmentation of the scene is carried out using a projector, with the advantage of rendering the augmentation visible to all onlookers simultaneously without calling for any to hold a handheld device such as a tablet or for each to wear some form of head-mounted display. In carrying out spatial AR, however, care must be taken to appropriately warp an image to be projected in a manner that it, when projected, appear free of distortions to the viewer. For planar scene geometry (such as a floor, wall, or table), a manual process called keystone correction can be used to carry out an appropriate corrective image warp, a process that is cumbersome and time consuming. Another drawback of conventional projector-based spatial AR is that it is capable of augmenting only the portion of the scene within the projector's own static field of view, thereby hindering its applicability to use cases calling for augmentation of wide areas such as a factory floorspace.

We propose a spatial AR system for wide-area metric augmentation of planar scene surfaces that produces the effect of keystone correction analytically as a function of the relative geometry of the projector and scene plane, using a projector equipped with a steerable mirror to direct the projection across varying target locations and a camera facing the scene plane in support of calibrating the system. Our system renders the placement of augmentations in the scene more intuitive than manual keystone correction in two ways. First, (i) the horizontal and vertical axes of the desired augmentations is carried out in accordance with the horizontal and vertical image axes of the camera, thereby making setting those axes as simple as appropriately rotating the camera relative to the scene. Second, (ii) the desired dimensions of the projected augmentations are specified in metric terms, thereby allowing for consistent scaling across all target locations. %Our system produces such augmentations accurately, with compelling time savings relative manual keystone correction.
\end{abstract}

@@

Calibration of a projector (or camera) in the sense we employ the term here\footnote{We are referring to a geometric calibration; not, e.g., to a color calibration.} renders one able to project a scene point~$\mathbf{X} \in \mathbb{R}^3$ to its corresponding pixel~$\mathbf{x} \in \mathbb{R}^2$ in the projector's (or camera's) image plane, or to compute the `back-projection' of $\mathbf{x}$, i.e., the ray from the projector's (or camera's) center of projection through $\mathbf{x}$ along which a projecting point~$\mathbf{X}$ must lie. Such a calibration can be expressed in terms of (i) a $3~\times{}~3$ calibration matrix~$\mathtt{K}$ derived from the projector's (or camera's) focal length~$f$ and principal point~$\mathbf{p}_0 = (x_0, y_0)^\top \in \mathbb{R}^2$ \cite{Hartley2004}, and (ii)~the coefficients of a lens distortion model used to correct for radial or tangential distortions caused by the lens system \cite{duane1971close,weng1992camera}; the calibration matrix~$\mathtt{K}$ is then
\begin{equation}
\mathtt{K} = \begin{bmatrix}
f & 0 & x_0 \\
0 & f & y_0 \\
0 & 0 & 1
\end{bmatrix},
\end{equation}
where $f$ and $(x_0, y_0)^\top$ are both expressed in units of pixels. The calibration matrix~$\mathtt{K}$ models a so-called pinhole camera; such a model is applicable to an image acquired using a real-world camera assuming the image has been corrected for lens distortion effects, by having applied the appropriate lens distortion model coefficients.

@@

\begin{figure}
    \centering
    \subfloat[\centering Asymmetrical circle pattern image, in image plane of projector (detections overlaid).]{{\includegraphics[height=3.25cm]{images/circles.png} }}
    \qquad
    \subfloat[\centering Projector calibration image (one for each target location), in image plane of camera (detections overlaid).]{{\includegraphics[height=3.25cm]{images/detections.png}}}
    \caption{Recovering 2D positions in support of projector calibration. (a) The asymmetrical circle pattern that we project to each of the target locations in the scene plane. 2D positions of the 2D-3D correspondences to be used for calibrating the projector are detected automatically, as the respective circle centers in the asymmetrical circle pattern image and given in a fixed ordering. (b) For each such target location, a projector calibration image is acquired from the viewpoint of the camera and the circle centers of the projected asymmetrical circle pattern are detected using the same algorithm as in (a). A chessboard pattern to be used for recovering the local scene plane is placed near the projected pattern, whose corners are likewise detected. Detected 2D projected asymmetrical circle pattern circle center points and chessboard corners overlain for illustration.}
    \label{fig:2d}
\end{figure}

\paragraph{2D-3D correspondences for camera calibration} We recover 2D-3D correspondences in support of calibrating the downwards-facing camera by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners ($8~\times{}~7$ squares), with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units in terms of which the camera calibration is carried out, and---owing to how our projector calibration relies on the camera calibration---the units of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting be off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$.

\paragraph{2D-3D correspondences for projector calibration} As in camera calibration, we calibrate the projector by relying on 2D-3D correspondences. For projector calibration, however, obtaining those correspondences involves \textit{projecting} a calibration pattern to the ground plane. The 2D pixel positions of the projected pattern are obtained by

 are known \textit{a priori}      that we detect using the calibrated downwards-facing camera.

 The pattern we project is one of circles (cf.\ Figure~\ref{fig:2d}(a)), and we rely on an algorithm to detect the circle center points in the asymmetrical circle pattern in the image plane of the projector \cite{bradski2000opencv}, giving the 2D positions of our 2D-3D correspondences for calibrating the projector. We project the asymmetrical circle pattern to each of the target locations, and use the the calibrated downwards-facing camera to acquire a projector calibration image for each. Given a projector calibration image acquired using camera, we detect the circle centers of the \textit{projected} asymmetrical circle pattern (cf.\ Figure~\ref{fig:2d}(b)); given the scene plane (the recovery of which we shall return to in the paragraph that follows) and such a 2D circle center $\mathbf{x}$, we obtain its 3D correspondence by intersecting the back-projection~$\mathtt{K}_\text{cam}^{-1}(\mathbf{x}^\top, 1)^\top \in \mathbb{P}^2$ of $\mathbf{x}$ with the recovered scene plane (cf.\ Figure~\ref{fig:3d}). Since the algorithm that yields 2D circle centers does so in a consistent ordering, we thus obtain the 2D-3D correspondences between the projector's image plane and the scene required for projector calibration, yielding the projector calibration matrix~$\mathtt{K}_\text{proj}$.

\begin{figure}
    \centerline{\includegraphics[scale=.375]{images/2d3d.png}}
    \caption{Obtaining 2D-3D correspondences for projector calibration, using a projector calibration image (cf.\ Figure~\ref{fig:2d}(b)). First, the scene plane (gray) is recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circle center points of the asymmetrical circle pattern---i.e., the 3D positions of the 2D-3D correspondences to be used for calibrating the projector (red)---obtained by intersection with the scene plane of back-projections (likewise gray) of the 2D circle center points of the asymmetrical circle pattern detected in the image plane. The downwards-facing camera is shown in terms of its camera frustum (black, with up vector in red).}
    \label{fig:3d}
\end{figure}

@@

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/3_cam.png}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_cam.png}}}
    \caption{Obtaining the virtual camera. A physical camera placed to face downwards is almost certain to not face downwards precisely. We correct for this possibility by rotating the recovered downwards-facing camera (black) about its center of projection such that its optical axis be made parallel with the plane's normal. Since it is in accordance with our downwards-facing camera the we wish to orient our virtual projector and since our virtual projector is required to face directly downwards to the scene plane to compute our corrective image warp, it is the orientiation of the virtual camera that we use to subsequently orient our virtual projector.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera. This rotated camera is our virtual camera, and it is the orientation of the virtual camera that we subsequently use to orient our virtual projector.
    \label{fig:virtualcam}
\end{figure}

\begin{figure}
    \centering
    \subfloat[\centering Oblique view (annotated).]{{\includegraphics[height=4.2cm]{images/1_big.pdf}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3_proj.png}}}
    \caption{Obtaining the virtual projector. The virtual projector is then obtained by (i) rotating the recovered projector (black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) translating along the normal direction to achieve the desired metric projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

@@

\paragraph{Plane-induced homography} Let $\mathtt{K}_\text{proj}$ express the calibration matrix of the recovered projector and $(\mathtt{R}, \mathbf{t}) \in SE(3)$ the rigid body transformation that transforms points from the coordinate frame of the recovered projector to that of the virtual projector, for a given target location. Moreover, let $(\mathbf{n}^\top, -d)^\top$ give the scene plane, expressed in the coordinate frame of the recovered projector, where $\mathbf{n} \in \mathbb{R}^3$ is the scene plane's normal vector and $d = \mathbf{n}^\top\mathbf{X}$ for any point~$\mathbf{X}$ in the plane, so that $(\mathbf{n}^\top, -d) (\mathbf{X}^\top, 1)^\top = 0$. The transformation that warps the image to be projected to the scene plane by the recovered projector such that it appear as if were projected to the scene plane by the virtual projector (cf.\ Figure~\ref{fig:warp}(b)) is given the by the $3~\times~3$ matrix
\begin{equation}
\mathtt{H} = \mathtt{K}_\text{proj}\left(\mathtt{R} - \frac{\mathbf{t}\mathbf{n}^\top}{d}\right)\mathtt{K}_\text{proj}^{-1},
\label{homgen}
\end{equation}
a form of `plane-induced' 2D homography \cite{Hartley2004}. For convenience, we enable optional rotation of the image to be projected \textit{before} applying $\mathtt{H}$, about the image center; that rotation, parameterized in degrees, is thus in effect carried out intuitively with respect to the ground plane.
